---
title: "PA1_template.Rmd"
author: "Salah Uddin"
date: "20 September 2015"
output: html_document
---

#Loading and preprocessing the data
```{r}

data<- read.csv(file("activity.csv","r"))

class(data[,2])

data[,2]<- as.Date(data[,2])
```
#What is mean total number of steps taken per day?
```{r}

t<-tapply(data$steps, as.Date(data$date, format="%m/%d/%Y"), sum)
perday<-t
i=1
while(i<=length(t)){
        perday[i]<- t[[i]];
        i=i+1;
}

library(ggplot2)

udate<-unique(data$date)

dat<-data.frame(perday,udate)



library(scales)
library(dplyr)


g<-ggplot(data=dat,aes(x=udate,y=perday))+ geom_bar(stat="identity") + scale_x_date(labels=date_format("%Y-%b"),breaks="1 day") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
print(g)

mean(perday,na.rm = T)
median(perday, na.rm = T)
```
#What is the average daily activity pattern?
```{r}

steps<-data$steps
interval<-data$interval
class(data$interval)
groupbyinterval <-aggregate(steps ~ interval, data = data, mean)

summary(groupbyinterval)


plot(x=groupbyinterval$interval,y=groupbyinterval$steps, type = "l")
apply(groupbyinterval, MARGIN = 2, function(x) max(x, na.rm=TRUE))

```
#Imputing missing values


##creating a new data set my strategy will be to fill it with the mean

```{r}


na_rows<-data[!complete.cases(data),]

non_na_rows<- data[complete.cases(data),]


na_rows_new <- mutate(na_rows, steps=round(mean(groupbyinterval$steps)))

newset<- rbind(na_rows_new,non_na_rows)


data<-newset

class(data[,2])

data[,2]<- as.Date(data[,2])

t<-tapply(data$steps, as.Date(data$date, format="%m/%d/%Y"), sum)




i=1
while(i<=length(t)){
        perday[i]<- t[[i]];
        i=i+1;
}

udate<-unique(data$date)

dat<-data.frame(perday,udate)





g<-ggplot(data=dat,aes(x=udate,y=perday))+ geom_bar(stat="identity") + scale_x_date(labels=date_format("%Y-%b"),breaks="1 day") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
print(g)
```
# Dont take the date as it is automated by ggplot
```{r}

mean(perday,na.rm = T)
median(perday, na.rm = T)
```
#Are there differences in activity patterns between weekdays and weekends?

Not much difference
```{r}

library(timeDate)
data$date<-as.Date(data$date)
wdate<-isWeekday(data$date, wday=1:5)

data<- data.frame(data,wdate)
```
#true means week days and false means weekend
```{r}

i=1
while(i<=length(wdate)){
        if(data$wdate[i]==T){
                data$wdate[i]<- "WeekDay"
                
                
        }else{
                data$wdate[i]<- "WeekEnd"
        }
          i=i+1      
        
}

data$wdate<- as.factor(data$wdate)

steps<-data$steps
interval<-data$interval

winterval <-aggregate(steps ~ interval + wdate, data = data, mean)
library(lattice)
xyplot(winterval$steps ~ winterval$interval|winterval$wdate, winterval, type = "l", layout = c(1, 2))

```
